% !TEX root = ../Planning.tex

\section{Documentation (R-D)}
\label{sec:documentation}
\subsection{Documentation plan}
Just like the project itself, the documentation will be managed as an iterative product.
All project documentation templates, forming the basis for the documentation deliverables, will be created before the start of phase 3b, investigation context, and all analysis, design and coding documentation will be maintained on a weekly basis.

Each sprint delivery, the available set of documentation will be delivered informally towards the project supervisor and, where useful, towards the client.
The project team strives to finalize all intermediate documentation at least one week after the release corresponding deliverable.

\subsection{Documentation deliverables}
All documentation deliverables, listed in section `Project planning, milestones and corresponding deliverables', are explained in some more detail in this section.

\subsubsection{Detailed project plan}
The detailed project plan, this document, covers at least the following topics:

	\begin{enumerate}
		\item Project description, context and goals
		\item Project approach and project management
		\item Planning
		\item Risks and risk management approach
		\item Quality approach and control
		\item Project requirements, global architecture and the main component, including the environment, of the to be modified system
		\item Test plan
		\item Used tools, methodologies and documentation
		\item knowledge gathering for proper project execution
		\item Communication plan
		\item Glossary
	\end {enumerate}

\subsubsection{Detailed allocation of tasks within the project team}
This project will be driven by 2 project team members who will work closely together.
Based on the personal selection of the  academic competencies, some highlights, responsibilities and corresponding roles will be assigned to each individual team member.

This document will clarify the assigned roles and tasks each team member will take care of.

The `detailed allocation of tasks within the project team' is integrated in the deliverable `Detailed project plan'.

\subsubsection{Scientific article}
Each individual project team member will deliver an in depth analysis of one or more topics regarding the environment in which the system is, or will be, used.

This deliverable will cover at least the following topics:

	\begin{enumerate}
		\item Problem definition
		\item Motivation
		\item Context
		\item Used analysis techniques
		\item Results
		\item Conclusions
		\item Glossary
		\item Bibliography
		\item Appendices, if useful
	\end {enumerate}

\subsubsection{Detailed report of the `investigation context'}
The detailed report of the `investigation context' will cover at least following topics:

	\begin{enumerate}
		\item Context
		\item Conclusions of the researcher
		\item Conclusions in regards to the project
		\item Glossary
		\item Bibliography
		\item Appendices, if useful
	\end {enumerate}

\subsubsection{Analysis \& design}
The `analysis \& design document' will cover at least following topics:

	\begin{enumerate}
		\item Design decisions and the reasoning behind them
		\item Detailed solution design
		\item Implementation guidelines \& coding conventions
		\item Tools \& libraries used in the project
	\end {enumerate}

\subsubsection{Test report}
The `Test report' will cover at least the following topics:

	\begin{enumerate}
		\item Overview of test scenario's
		\item Test results
		\item Conclusions
	\end {enumerate}

\subsubsection{Source code}
The source code will be document in the code itself.
It is of utmost importance that the code is documented in a clear and concise way.
Topics needing a more elaborated documentation approach are documented in the deliverable IT documentation.

\subsubsection{IT documentation}
Some coding structures, or the architectural design principles behind it, can require some additional documentation which could pollute the source code itself.
The IT documentation deliverable will list all complex coding structures and the clarification of them including the overall technical architecture, in accordance with the project scope, and the technical design principles.

The IT documentation must provide the necessary support to our customer to fully understand the project coding and it needs to present a solid basis on which future design and implementation decisions can be taken.

\subsubsection{Project documentation document}
The project will be concluded in a general project documentation document.
This deliverable will conclude all main deliverables, where possible, in a brief way.

Following additional topics will be documented to close the project:
	\begin{enumerate}
		\item Main realizations
		\item Encountered issues \& risks
		\item Tools \& methodologies effectiveness
		\item Topics not delivered with the project
		\item Budget analysis
		\item Lessons learned
		\item Conclusions
	\end {enumerate}