% !TEX root = ../Thesis.tex

\subsection{Lessons learned}
\label{assessment:lessons-learned}
This paragraph summarizes our lessons learned, both the positive ones and the ones in which we see some room for improvement.

  \begin{description}
    \item[Planning] Delivering a project like this one after working hours was hard for both participants. 
	  Both Daniel and Maarten have full-time jobs.
	  In our planning, we overestimated our availabilities, due to which we had some issue to adhere to our initial project plan.
	  Next time, we would foresee more personal time besides the project work.
	
    \item[Customer alignment] is very important to correctly understand the project goals and expectations.
	  There were several alignment sessions with the customer, but due to our time constraints, we focused a lot on the delivery of the project. 
	  Next time we would plan even more alignments on a regular basis with the customer.	  
    
    \item[Proper analysis] is crucial in a project like this one.
	  During the project, we realized that the result of choosing an inappropriate parser would have been catastrophic.
	  Given the obliged project structure, we luckily took the time to define the correct parser.
      Should we redo this project, we would reserve the same amount of time to do this analysis.
    
    \item[Teamwork] makes the difference.
	  It does not matter how hard a project is, if the team keeps on supporting each other, results will be made.
	  Daniel and Maarten took the time to get to know each other.
	  This resulted in a high mutual understanding and a great deal of respect.
	  Although the project was hard by times, this respect made the difference and both team members supported each other where possible.
	  The project would not have this result without this nice piece of excellent teamwork.
	 
	 \item[The expertise of the project supervisor] will lift a team to a higher level.
	  The support we received from our supervisor was priceless.
	  Each time the project team experienced an issue or had some questions, regardless the context, Bastiaan showed us the right direction.
	  Without his support, the progress and quality of the project would be of a much lower level.
  \end{description}
