% !TEX root = ../Documentation.tex

\subsection{Errors}
\label{subsec:design-errors}
The proof of the pudding is in the eating, and therefore, our error message qualification definition from \autoref{subsec:analysis-errors} is applied on the new parser.
All erroneous syntax statements recorded during the analysis phase are fed to the new parser to draw up the error message quality overview of the new parser.

The results are placed in \autoref{tab:error-messages-results} next to the results of the old parser as a reference point.
The actual error list, containing the syntax statements with the corresponding error messages, is available in \autoref{app:error-list}.

% Please add the following required packages to your document preamble:
% \usepackage[table,xcdraw]{xcolor}
% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
\begin{table}[h]
  \centering
	\begin{tabular}{llrlr}
    Error quality  & \multicolumn{2}{c}{Old parser} & \multicolumn{2}{c}{New parser} \\
		Good           & 19          & 22,35\%          & 70          & 82,35\%          \\
		Average        & 18          & 21,18\%          & 14          & 16,47\%          \\
		Bad            & 48          & 56,47\%          & 1           &  1,18\%          \\
		\rowcolor[HTML]{BBBBBB}
		\textbf{Total} & \textbf{85} & \textbf{100,00\%} & \textbf{85} & \textbf{100,00\%}
	\end{tabular}
  \caption{Error message comparison results}
  \label{tab:error-messages-results}
\end{table}

The table clearly visualizes that there indeed was an issue with the error messages generated by the old Ampersand parser.
By implementing the new parser, an improvement of 360\% is made towards the creation of good error messages in which 82\% of the generated error messages are good, compared to 22\% in the old Parser.

One bad error message remains in the system.
After analysis, the resolution of this remaining bad error would require too much effort, hence increased complexity, compared to the actual value gain.
The error in question is thrown when the \hyperref[fig:ebnf-Meta]{\texttt{Meta}} statement is used with only one string value. 
Given the simple structure of the notation, we assume that the reference to the line number of the \hyperref[fig:ebnf-Meta]{\texttt{Meta}} is sufficient although the error message is unclear.
Therefore, this message is kept as is.
